Model-agnostic Interpretable Machine Learning

### 2021-08-19 2021-bips-practice-phd-defence/2021-bips-practice-phd-defence.Rmd
add ice and pdp explanations

## Some ideas

- Cite pitfalls at many places in the talk
- INtroduction with History: INterweave motivation, and also cite history paper

@@

@@

## Individual Conditional Expectations (ICE)

maybe use ice + pdp animation, but split into individual frames so they are clickable

@@

The estimator:
$$\widehat{PD} = \frac{1}{n_2}\sum_{i=1}^{n_2} \hat{f}(x, x_{C}^{(i)})$$

What it looks like:

TODO: 1D-plot and 2D-plot

>>>>

@@

## Model-agnostic Interpretation

@@

## Individual Conditional Expectation (ICE)

How does a feature influence the prediction for an individual observation?

```{r}
load(url("https://github.com/christophM/interpretable-ml-book/blob/master/data/bike.RData?raw=true"))
library("mlr")
library("iml")
library("ggplot2")
library(dplyr)
library(tidyr)

bike.task = makeRegrTask(data = bike, target = "cnt")
mod.bike = mlr::train(mlr::makeLearner(cl = 'regr.randomForest', id = 'bike-rf'), bike.task)

set.seed(42)
bike.sub = bike[sample(1:nrow(bike), size = 15),]

pred.bike = Predictor$new(mod.bike, data = bike.sub)
pdp = FeatureEffect$new(pred.bike, "temp", method = "ice", grid.size = 7)

res.raw = pdp$results
theme_set(theme_classic(base_size = 20))

## Step 1: Define grid over feature
grid.dat = data.frame(temp = unique(res.raw$temp), .id = 1, temp2 = 1)
p = ggplot(grid.dat) + geom_vline(aes(xintercept = temp)) +
  scale_y_continuous("Prediction", limits = c(min(res.raw$.y.hat), max(res.raw$.y.hat))) +
  ggtitle("1) Define grid along feature") +
  scale_x_continuous("Feature")
  #scale_y_continuous("Prediction")
p
```

## Individual Conditional Expectation (ICE)

How does a feature influence the prediction for an individual observation?

```{r}
## Step 2: Get predictions for data set when at grid
grid.dat = data.frame(temp = unique(res.raw$temp))
res.raw$tt = res.raw$temp
p = ggplot(res.raw, aes(x = temp, y = .value)) +
  geom_vline(data = grid.dat, aes(xintercept = temp)) +
  geom_point(aes(color = factor(.id)), size = 3)+
  scale_color_discrete(guide = "none") +
  ggtitle("2) Model predictions at grid points") +
  scale_x_continuous("Feature") +
  scale_y_continuous("Prediction", breaks = NULL)
p
```

## Individual Conditional Expectation (ICE)

How does a feature influence the prediction for an individual observation?

```{r}
res.raw$.id2 = factor(res.raw$.id)
p = ggplot(res.raw, aes(x = temp, y = .value))  +
  #geom_vline(aes(xintercept = temp), data = grid.dat) +
  geom_line(aes(group = .id), size  = 1)  +
  geom_point(data = res.raw[c("temp", ".value", ".id2")], aes(color = .id2), size = 3) +
  scale_color_discrete(guide = "none") +
  ggtitle("3) Line per data instance -> ICE curves") +
  scale_x_continuous("Feature") +
  scale_y_continuous("Prediction", breaks = NULL)
p
```

How does a feature influence the prediction for an individual observation **on average**?

How does a feature influence the prediction for an individual observation **on average**?

```{r}
p
```

How does a feature influence the prediction for an individual observation **on average**?

```{r}
res = res.raw %>% group_by(temp) %>%
  summarize(pdp = mean(.value))

p = ggplot(res) +
  geom_line(aes(x = temp, y = pdp), size = 1.5) +
  geom_point(data = res.raw[c("temp", ".value", ".id2")], aes(color = .id2, x = temp, y = .value), size = 1, alpha = 0.5) +
  ggtitle("4) Average curves to get a PDP") +
  scale_color_discrete(guide = "none") +
  scale_x_continuous("Feature") +
  scale_y_continuous("Prediction", breaks = NULL)
p
```

@@

The estimator:
$$\widehat{PD} = \frac{1}{n}\sum_{i=1}^{n} \hat{f}(x, x_{C}^{(i)})$$
